---
title: "Simulation power + FDR/FWER of novel method"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(purrr)
library(ggplot2)
library(reshape2)
library(patchwork)
library(Rcpp)
library(RcppArmadillo)
library(zeallot)
library(repfdr)
# remotes::install_github("jingshuw/adaFilter")
library(adaFilter)
# BiocManager::install("limma")
library(limma)
# remotes::install_github("https://github.com/jdreyf/DirectionalMaxPTest/tree/4d94d1be714e8ce193928412354e766d9147da42")
library(DirectionalMaxPTest)
# remotes::install_github("shay-y/radjust")
library(radjust)

source("R/mvrnormArma_rcpp.R")
source("R/safe_repfdr.R")
source("R/sim_mu_vctrs.R")
source("R/SimFDR.R")
source("R/SimFWER.R")
source("R/sim_ggplot.R")
```

# Simulations

We simulate ScreenDMT vs. competitors in different settings to compare false positive rates and power. These simulations are analogous to Vera's Biometrical Journal paper Figure 3.

```{r 4sim}
# i want to focus on low pi1, since that's common
pi1.v <- c(0, 0.01*2**(0:5))
pi2 <- 0.05
# i get repfdr error that pi0theo is missing when m <= 200
# m.v <- c(200, 10**3)
m <- m.v <- 1000
fac.v <- 1:2
rho <- 0.5
nblock.v <- c(10, 100)
alpha <- 0.05
B <- 10**3
mu <- 3.5
```

## Sim FDR 

```{r simfdr, include=FALSE, eval=TRUE}
meth.v <- c("Pearson", "DMT", "ScreenDMT", "radjust_sym", "AF", "repfdr")
type.v <- c("FDR", "power")

pwr_fdr_arr.dnms <- list(paste0("pi1_", pi1.v), paste0("m_", m.v), paste0("fac_", fac.v), type.v, meth.v, paste0("nb_", nblock.v))
pwr_fdr_arr <- array(dim = sapply(pwr_fdr_arr.dnms, length), dimnames = pwr_fdr_arr.dnms)

print(Sys.time())
for (nb in nblock.v){
  for (pi1 in pi1.v){
    pi0 <- 1 - pi2 - pi1
    for (m in m.v){
      for (fac in fac.v){
        pwr_fdr_arr[paste0("pi1_", pi1), paste0("m_", m), paste0("fac_", fac), type.v, meth.v, paste0("nb_", nb)] <- 
        SimFDR(mu = mu, pi0 = pi0, pi1 = pi1, pi2 = pi2, alpha = alpha, m = m, B = B, fac=fac, nblocks = nb, rho = rho, meth.v=meth.v)
        
        stopifnot(!is.na(pwr_fdr_arr[paste0("pi1_", pi1), paste0("m_", m), paste0("fac_", fac), type.v, meth.v, paste0("nb_", nb)]))
      }
      cat("\n m:", m, "done \n\n")
    }
    cat("\n pi1:", pi1, "done \n\n")
  }
  cat("\n nb:", nb, "done \n\n")
  print(Sys.time())
}

stopifnot(length(m.v) == 1)
saveRDS(pwr_fdr_arr, file=paste0("results/pwr_fdr_arr_b", B, "_m", m.v, "_mu", mu, "_logpi1.RDS"))
```

We simulate the various settings and capture power and FDR in multi-dimensional array, where each dimension is a parameter we vary e.g. "pi1", stored at `r paste0("results/pwr_fdr_arr_b", B, "_m", m.v, "_mu", mu, "_logpi1.RDS")`.

## Sim FWER 

```{r simfwer, include=FALSE, eval=TRUE}
meth.v <- c("Pearson", "DMT", "ScreenDMT", "AF")
type.v <- c("FWER", "power")

pwr_fwer_arr.dnms <- list(paste0("pi1_", pi1.v), paste0("m_", m.v), paste0("fac_", fac.v), type.v, meth.v, paste0("nb_", nblock.v))
pwr_fwer_arr <- array(dim = sapply(pwr_fwer_arr.dnms, length), dimnames = pwr_fwer_arr.dnms)

print(Sys.time())
for (nb in nblock.v){
  for (pi1 in pi1.v){
    pi0 <- 1 - pi2 - pi1
    for (m in m.v){
      for (fac in fac.v){
        pwr_fwer_arr[paste0("pi1_", pi1), paste0("m_", m), paste0("fac_", fac), type.v, meth.v, paste0("nb_", nb)] <- 
        SimFWER(mu = mu, pi0 = pi0, pi1 = pi1, pi2 = pi2, alpha = alpha, m = m, B = B, fac=fac, nblocks = nb, rho = rho, meth.v=meth.v)
      }
      cat("\n m:", m, "done \n\n")
    }
    cat("\n pi1:", pi1, "done \n\n")
  }
  cat("\n nb:", nb, "done \n\n")
  print(Sys.time())
}

stopifnot(length(m.v) == 1)
saveRDS(pwr_fwer_arr, file=paste0("results/pwr_fwer_arr_b", B, "_m", m,  "_mu", mu, "_logpi1.RDS"))
```

We simulate the various settings and capture power and FWER in `r paste0("results/pwr_fwer_arr_b", B, "_m", m,  "_mu", mu, "_logpi1.RDS")`.

# Plots

```{r ggplot}
for (type1 in c("fdr", "fwer")){
  pwr_arr <- readRDS(paste0("results/pwr_", type1, "_arr_b", B, "_m", m, "_mu", mu, "_logpi1.RDS"))
  ggp.lst <- sim_ggplot(pwr_arr)
  ggp2 <- ggp.lst[[1]] / ggp.lst[[2]]
  ggsave(paste0("results/ggsim_b", B, "_m", m, "_mu", mu, "_", type1, ".pdf"), plot = ggp2, width=7, height=10)
}
```

We plot the power and FDR at `r paste0("results/ggsim_b", B, "_m", m, "_mu", mu, "_fdr.pdf")` and power and FWER at `r paste0("results/ggsim_b", B, "_m", m, "_mu", mu, "_fwer.pdf")`.

# R session
```{r, include=TRUE}
pander::pander(sessionInfo())
```